On hyperparameter tuning through Lipschitz global optimization

### 2024-03-07 2023/ITMO paper/elsarticle-template-num.tex
reformulate theorem in Russian

Let us number all the different combinations of categorical parameters with integers $s, 1\leq s \leq S$, i.e., let us assign a vector $u_s$ to each number $s$. Then the problem under consideration can be written in the form
\begin{gather}\label{problem_is}
 \min_{s\in\{1,...,S\}}\left[\min{\left\{ \varphi(y,u_s) : y\in D \right\}}\right].
\end{gather}

Using a dimensionality reduction scheme with Peano curves $y(x), x\in [0,1]$,  we can assign to each minimization problem on $y$ a one-dimensional minimization problem
\[
 \min{\left\{ \varphi(y(x),u_s): x \in [0,1] \right\}}, s \in \{1,...,S\}.
\]

@@

Условия сходимости Global Search Algorithm для задач с Categorical Parameters (GSA-CP)  могут быть сформулированы в виде следующей теоремы.

\textbf{Theorem} (convergence conditions)

Пусть в постановке задачи (\ref{problem_is1}) функция $f(x)$ из (\ref{f}) удовлетворяет условию Гельдера с константой $H$ при $x \in (i, i + 1), 0 \leq i < S$.
И пусть точка $\overline{x}$ be a limit point of the sequence $\left\{x^k\right\} $, полученной при минимизации функции $f(x)$ с помощью алгоритма GSA-CP. Then:
\begin{enumerate}
	\item 
	если одноверменно с $\overline{x}$ существует другая предельная точка $x'$ последовательности $\left\{x^k\right\}$, тогда $f(\overline{x}) = f(x')$.
	\item
	$z^k=f(x^k) \geq f(\overline{x})$ при любом $k \geq 0$.
	\item
	Если на некотором шаге поиска величина $\mu$ from (\ref{mu}) satisfies the condition
\begin{equation}\label{217}
  r \mu > 4H,
\end{equation}
then $\overline{x}$ is a global minimizer of the function $f(x), x\in[0,S]$ and any global minimizer $x^\ast$ is also a limit point of the sequence $\left\{x^k\right\}$.
\end{enumerate}

\textbf{Proof}. Эта теорема может быть доказана аналогично теореме о сходимости index method of global optimization from \cite{Strongin2000}, принимая во внимание формулу (\ref{f}), определяющую функцию $f(x)$, правила (\ref{index}), задающие индексы у точек испытаний, а также взаимосвясь (\ref{holder}) между Lipschitz and H\"older conditions при редукции размерности с помощью evolvents.

>>>>

Let us number all the different combinations of categorical parameters with integers $s = 1,...,S$, i.e., let us assign a vector $u_s$ to each number $s$. Then the problem under consideration can be written in the form
\begin{gather}\label{problem_is}
 \min_{s\in\{1,...,S\}}\left[\min{\left\{ \varphi(y,u_s) : y\in D \right\}}\right].
\end{gather}

Using a dimensionality reduction scheme with Peano curves $y(x), x\in [0,1]$,  we can assign to each minimization problem on $y$ a one-dimensional minimization problem
\begin{equation}\label{fi_s}
 \min{\left\{ \varphi(y(x),u_s): x \in [0,1] \right\}}, s = 1,...,S.
\end{equation}

@@

%Новый текст на русском
\textcolor[rgb]{1,0,0}{Условия сходимости Global Search Algorithm для задач с Categorical Parameters (GSA-CP) могут быть сформулированы в виде следующего утверждения.}

\textbf{Утверждение} (сonvergence conditions)

Пусть функции $ \varphi(y,u_s), y \in D, s = 1,...,S,$ form (\ref{problem_is}) удовлетворяют условию Липшица с соответствующими константами $L_s, s = 1,...,S$, и $L = \max\{L_s\}$.
И пусть точка $\overline{x}$ be a limit point of the sequence $\left\{x^k\right\} $, полученной при минимизации функции $f(x)$ из (\ref{f}) с помощью алгоритма GSA-CP. Then:
\begin{enumerate}
	\item 
	если одноверменно с $\overline{x}$ существует другая предельная точка $x'$ последовательности $\left\{x^k\right\}$, тогда $f(\overline{x}) = f(x')$.
	\item
	$z^k=f(x^k) \geq f(\overline{x})$ при любом $k \geq 0$.
	\item
	Если на некотором шаге поиска величина $\mu$ from (\ref{mu}) satisfies the condition
\begin{equation}\label{217}
  r \mu > 2^{3-1/N} L \sqrt{N+3},
\end{equation}
then $\overline{x}$ is a global minimizer of the function $f(x), x\in[0,S]$ and any global minimizer $x^\ast$ is also a limit point of the sequence $\left\{x^k\right\}$.
\end{enumerate}

\textbf{Proof}. Это утверждение доказывается аналогично теореме о сходимости index method of global optimization from \cite{Strongin2000}, принимая во внимание взаимосвясь (\ref{f}) функций  $f(x)$ и $\varphi(y,u_s)$, а также правила (\ref{index}), задающие индексы у точек испытаний.
